---
title: "Bryan_Lee_Computer_HW_4"
author: "Bryan Lee"
date: "2024-05-16"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

#Exercise 1: A company with a large fleet of cars wants to study the gasoline usage. They check the gasoline usage for 50 company trips chosen at random, finding a mean of 25.16 mpg. Based on the past experience, they believe that the gasoline usage is normally distributed and the standard deviation of the general gasoline usage is 4.82 mpg

#(a) Which kind of confidence intervals is appropriate to use here, z-interval or t-interval?
#z-interval as standard deviation is known and n>40.

#(b) Please use R to find the critical value they need when constructing a 90% CI.
```{r}
n = 50
sd = 4.82
x_bar = 25.16

qnorm(1-((1-0.90)/2))
```
#(c) Please use R to construct a 90% CI for the mean of the general gasoline usage.
```{r}
n = 50
sd = 4.82
x_bar = 25.16

#Lower bound
x_bar-qnorm(1-((1-0.90)/2)) * (sd/sqrt(n))

#Upper bound
x_bar+qnorm(1-((1-0.90)/2)) * (sd/sqrt(n))
```

#(d) If they want to control the width of CI to be within 0.60 mpg, at least how many trips do they have to sample? Use R to calculate.
```{r}
w = 0.60
sd = 4.82
n = ((2*qnorm(1-((1-0.90)/2)) * sd)/w)^2
ceiling(n) #For rounding to the next int
```

#(e) Create a R function whose argument is the width of CI, and the output is the sample size necessary to achieve such accuracy. Confidence level is fixed at 90%.
```{r}
sample_size_for_CI = function(width) {
  CI = 0.9
  sd = 4.82
  
  z_crit = qnorm(1-((1-CI)/2))
  n = ((2 * z_crit * sd)/width)^2
  return(ceiling(n))
}

```

#(f) Apply the function you created in part (e) to demonstrate that larger sample size is required to achieve better accuracy (i.e, narrower CI width). Confidence level is fixed at 90%.
```{r}
sd = 4.82
x_bar = 25.16

#small sample & wider CI width
n = sample_size_for_CI(0.6)
n
#Lower bound
x_bar-qnorm(1-((1-0.90)/2)) * (sd/sqrt(n))

#Upper bound
x_bar+qnorm(1-((1-0.90)/2)) * (sd/sqrt(n))

#large sample & narrower CI width
n = sample_size_for_CI(0.1)
n
#Lower bound
x_bar-qnorm(1-((1-0.90)/2)) * (sd/sqrt(n))

#Upper bound
x_bar+qnorm(1-((1-0.90)/2)) * (sd/sqrt(n))
```

#Exercise 2: In a class survey, students are asked how many hours they sleep per night. In the sample of 21 students, the mean was 5.87 hours with a standard deviation of 1.56 hours. The parameter of interest is the mean number of hours slept per night in the population from which this sample was drawn.

#(a) Which kind of confidence intervals is appropriate to use here, z-interval or t-interval?
#z-interval as standard deviation is known.

#(b) Please use R to find the critical value they need when constructing a 98% CI.
```{r}
qnorm(1-((1-0.98)/2))
```

#(c) Please use R to construct a 98% CI for the mean number of hours slept per night.
```{r}
n = 21
x_bar = 5.87
sd = 1.56

#Lower bound
x_bar-qnorm(1-((1-0.98)/2)) * (sd/sqrt(n))

#Upper bound
x_bar+qnorm(1-((1-0.98)/2)) * (sd/sqrt(n))
```

#Exercise 3: In the year 2001 Youth Risk Behavior survey done by the U.S. Centers for Disease Control, 750 out of 1200 female 12th graders said they always use a seatbelt when driving.

a) We want to construct confidence interval to estimate the proportion of 12th grade females in the U.S. population who always use a seatbelt when driving. Is it it appropriate to use traditional confidence interval? Why or why not?
#It is appropriate to use traditional confidence interval because this situation is case 2 where standard deviation is unknown, n > 40, and needed to calculate population proportion.

#b) If answer ”Yes” in part (a), use R to find the 99% traditional CI for the proportion of 12th grade females in the population who always use a seatbelt when driving.
```{r}
n = 1200
p_hat = 750/1200

#Lower bound
p_hat-qnorm(1-((1-0.99)/2)) * sqrt((p_hat*(1-p_hat))/n)

#Upper bound
p_hat+qnorm(1-((1-0.99)/2)) * sqrt((p_hat*(1-p_hat))/n)
```
#(c) Assuming there is no prior information or past experience available, what is the sample size necessary to control the traditional 99% CI width to be within 0.01.
```{r}
p_hat = 1/2 #no prior info so 1/2 as per notes.
z = qnorm(1-((1-0.99)/2))
n = (4 * (z^2) * p_hat * (1-p_hat))/(0.01^2)
ceiling(n)
```

#Exercise 4: Consider the problem in Exercise 1 again. The company wants to conduct a test, with α = 0.05, to see whether the fleet average is less than 26 mpg.

#(a) Which kind of tests is appropriate to use here, z-test or t-test?
#z-interval as standard deviation is known and n>40. (exercise 1)

#(b) Write appropriate hypotheses. Is the alternative hypothesis one-sided or two-sided?
#We want to test mu = 26 where the alternative hypothesis to test is mu < 26 (one-sided).

#(c) Use R to compute the test statistic and construct rejection region. Make conclusion using the test statistic and rejection region.
```{r}
test_stat = (25.16 - 26)/(4.82/sqrt(50))
rej_region = c(-Inf, qnorm(0.05))

test_stat
rej_region

test_stat < qnorm(0.05)
#Conclusion: False, so test_stat does not lie within the rejection region, so we retain H0. 
```

(d) Use R to compute p-value. Make conclusion using p-value. Is it consistent with the conclusion in part (c)?
```{r}
alpha = 0.05
p_val = pnorm(test_stat)
p_val

p_val < alpha
#Conclusion: False, so retain H0, which is consistent with the conclusion from part c.
```
