---
  title: "MATH 324 Computer HW 3"
  author: "Bryan Lee"
  date: "May 7 2024"
  output:
  pdf_document: default
---

#Exercise 1: Based on weighting the most animals in the population, the American Angus Association reported that mature Angus cows had a mean weight of 1350 pounds with a standard deviation of 163 pounds. A research randomly measures 100 mature Angus cows.Use R commands to answers the following questions.

#(a) Find the probability that the average weight of the 100 mature Angus cows is less than 1325 pounds.
# P(X_bar < 1325) = pnorm(1325, mean, sd) where n = 100 and sd = 163/sqrt(100)
```{r}
pnorm(1325, 1350, 16.3)
```

#(b) Find the probability that the average weight of the 100 mature Angus cows is between 1325 and 1375
# P(1325 < X_bar < 1375)
```{r}
pnorm(1375, 1350, 16.3) - pnorm(1325, 1350, 16.3)
```

#Exercise 2: The lifetime of a special type of battery is a random variable with mean 40 and standard deviation 20 hours. Assume a stockpile of 50 independent such batteries.Use R commands to answer the following questions.

#(a) What is the probability that the total lifetime of the 50 batteries is less than 1900 hours?
# P(X_bar < 1900)
```{r}
mean = 50 * 40
s_d = sqrt(50) * 20
pnorm(1900, mean, s_d)
```

#(b) What is the probability that the total lifetime of the 50 batteries is more than 2200 hours?
# P(X_bar > 2200) = 1 - P(X_bar <= 2200)
```{r}
mean = 50 * 40
s_d = sqrt(50) * 20
1-pnorm(2200, mean, s_d)
```

#Exercise 3: The life time of a laser (in hours) is exponentially distributed with λ = 1/80. Two such lasers are operating independently. Use R commands to answer the following questions.

#(a) Use a simulated sample of size 1000 to estimate the probability that the sum of the two lifetimes is greater than 100 hours.
```{r}
n = 1000
lambda = 1/80
lifetimeA = rexp(1000, lambda)
lifetimeB = rexp(1000, lambda)

sum_lifetimes = lifetimeA + lifetimeB #sum of the lifetime

prob_sum_greater_than_100 = length(which(sum_lifetimes > 100)) / n #get all the lifetimes greater than 100 hrs and divide it by n to get prob
prob_sum_greater_than_100
```

#(b) Estimate the probability that the both lasers last less than 50 hours.
```{r}
n = 1000
lambda = 1/80
count = 0

for (i in 1:n) {
  if (lifetimeA[i] < 50 & lifetimeB[i] < 50) { #if both is less than 50 hrs
      count = count + 1
  }
}
count/n #The estimated prob (empirical), should be close to theoretical one.
(pexp(50, lambda))*(pexp(50, lambda)) # Theoretical probability that both lasers last less than 50 hours
```

#Exercise 4: Generate 500 samples, each of which is of size 5 and 50 (n=5, 50), from the following distributions:
#I. Uniform distribution on interval (0,5)
#II. Binomial distribution with n = 15, p = 0.2
#III. Exponential distribution with λ = 5
#IV. Poisson distribution with parameter µ = 2

```{r}
#Uniform distribution  on interval (0,5)
#Sample size 5:
unif_x_bar_list5 = rep(0,500)
unif_x_bar_list50 = rep(0,500)

for (i in 1:500) {
  sample5 = runif(5,0,5)
  unif_x_bar_list5[i] = mean(sample5)
}
mean(unif_x_bar_list5) #empirical mean (from sampling)
expectedValue = (5 + 0)/2 #theoretical expected (a+b)/2
expectedValue
var(unif_x_bar_list5) #empirical variance
sigma2 = (5 - 0)^2/12 #theoretical variance for each X_i (b-a)^2/12
sigma2/5 #theoretical variance for xbar. Should be similar to empirical variance

#Sample size 50:
for (i in 1:500) {
  sample50 = runif(50,0,5)
  unif_x_bar_list50[i] = mean(sample50)
}
mean(unif_x_bar_list50) #empirical mean
expectedValue = (5 + 0)/2 #theoretical expected (a+b)/2
expectedValue
var(unif_x_bar_list50) #empirical variance
sigma2 = (5 - 0)^2/12 #theoretical variance for each X_i
sigma2/50 #theoretical variance for xbar. Should be similar to empirical variance

hist(unif_x_bar_list5)
qqnorm(unif_x_bar_list5)
hist(unif_x_bar_list50)
qqnorm(unif_x_bar_list50)

#I've found that the empirical mean and variance of the x_bar list sampling of 5 and 50 are quite close to the theoretical mean and variance. A feature to be noted is that larger sample (i.e. sample size 50 compared to 5) will have a more symmetrical histogram. The sampling of the mean also follows the CLT as the histogram resembles a "bell curve" or a normal distribution, where with large sampling will create a stronger resemblance.
```

```{r}
#Binomial distribution with n = 15, p = 0.2
#Sample size 5:
n = 15
p = 0.2
binom_x_bar_list5 = rep(0,500)
binom_x_bar_list50 = rep(0,500)

for (i in 1:500) {
  sample5 = rbinom(5, n, p)
  binom_x_bar_list5[i] = mean(sample5)
}
mean(binom_x_bar_list5) #empirical mean (from sampling)
expectedValue = n*p #theoretical expected
expectedValue
var(binom_x_bar_list5) #empirical variance
sigma2 = n*p*(1-p) #theoretical variance for each X_i
sigma2/5 #theoretical variance for xbar. Should be similar to empirical variance

#Sample size 50:
for (i in 1:500) {
  sample50 = rbinom(50, n, p)
  binom_x_bar_list50[i] = mean(sample50)
}
mean(binom_x_bar_list50) #empirical mean (from sampling)
expectedValue = n*p #theoretical expected
expectedValue
var(binom_x_bar_list50) #empirical variance
sigma2 = n*p*(1-p) #theoretical variance for each X_i
sigma2/50 #theoretical variance for xbar. Should be similar to empirical variance

hist(binom_x_bar_list5)
qqnorm(binom_x_bar_list5)
hist(binom_x_bar_list50)
qqnorm(binom_x_bar_list50)

#I've found that the empirical mean and variance of the x_bar list sampling of 5 and 50 are quite close to the theoretical mean and variance. However, the empirical variance seems to have a bit of deviation from the theoretical value. A feature to be noted is that larger sample (i.e. sample size 50 compared to 5) will have a more symmetrical bell shaped histogram or a normal distribution (CLT), while a smaller sample will be more skewed shaped.
```

```{r}
#Exponential distribution with λ = 5
#Sample size 5:
lambda = 5
exp_x_bar_list5 = rep(0,500)
exp_x_bar_list50 = rep(0,500)

for (i in 1:500) {
  sample5 = rexp(5, lambda)
  exp_x_bar_list5[i] = mean(sample5)
}
mean(exp_x_bar_list5) #empirical mean (from sampling)
expectedValue = 1/lambda #theoretical expected
expectedValue
var(exp_x_bar_list5) #empirical variance
sigma2 = 1/lambda^2 #theoretical variance for each X_i
sigma2/5 #theoretical variance for xbar. Should be similar to empirical variance

#Sample size 50:
for (i in 1:500) {
  sample50 = rexp(50, lambda)
  exp_x_bar_list50[i] = mean(sample50)
}
mean(exp_x_bar_list50) #empirical mean (from sampling)
expectedValue = 1/lambda #theoretical expected
expectedValue
var(exp_x_bar_list50) #empirical variance
sigma2 = 1/lambda^2 #theoretical variance for each X_i
sigma2/50 #theoretical variance for xbar. Should be similar to empirical variance (8e-04 is 0.0008)

hist(exp_x_bar_list5)
qqnorm(exp_x_bar_list5)
hist(exp_x_bar_list50)
qqnorm(exp_x_bar_list50)

#I've found that the empirical mean and variance of the x_bar list sampling of 5 and 50 are quite close to the theoretical mean and variance. A feature to be noted is that larger sample (i.e. sample size 50 compared to 5) will have a histogram that resembles a bell shaped or a normal distribution (CLT), while a smaller sample will be skewed to the right.
```

```{r}
#Poisson distribution with parameter μ = 2
#Sample size 5:
mu = 2
pois_x_bar_list5 = rep(0,500)
pois_x_bar_list50 = rep(0,500)

for (i in 1:500) {
  sample5 = rpois(5, mu)
  pois_x_bar_list5[i] = mean(sample5)
}
mean(pois_x_bar_list5) #empirical mean (from sampling)
expectedValue = mu #theoretical expected
expectedValue
var(pois_x_bar_list5) #empirical variance
sigma2 = mu #theoretical variance for each X_i
sigma2/5 #theoretical variance for xbar. Should be similar to empirical variance

#Sample size 50:
for (i in 1:500) {
  sample50 = rpois(5, mu)
  pois_x_bar_list50[i] = mean(sample50)
}
mean(pois_x_bar_list50) #empirical mean (from sampling)
expectedValue = mu #theoretical expected
expectedValue
var(pois_x_bar_list50) #empirical variance
sigma2 = mu #theoretical variance for each X_i
sigma2/50 #theoretical variance for xbar. Should be similar to empirical variance

hist(pois_x_bar_list5)
qqnorm(pois_x_bar_list5)
hist(pois_x_bar_list50)
qqnorm(pois_x_bar_list50)

#I've found that the empirical mean and variance of the x_bar list sampling of 5 and 50 are quite close to the theoretical mean and variance. A feature to be noted is that larger sample (i.e. sample size 50 compared to 5) will have a histogram that resembles a bell shaped or a normal distribution (CLT), while a smaller sample will be a tad skewed with a tall peak. Although a larger sample still have a tall peek, it is more balanced out to resemble a bell shape.
```